\def\llncs{0}       % 1 = using llncs



\documentclass[11pt]{article}
%\usepackage[margin=1.25in]{geometry}
%\usepackage[letterpaper,hmargin=1.25in,vmargin=1.25in]{geometry}
\usepackage{fullpage}
\usepackage{amsthm}

\usepackage{amsthm,amsmath,amsfonts,amssymb,amstext}
\usepackage{latexsym,ifthen,url,rotating}

%\usepackage[pdftex,usenames,dvipsnames]{color}
\usepackage[usenames,dvipsnames]{color}

\ifnum\llncs=0
\ifx\pdfoutput\undefined 

%% latex

\else

%% pdflatex

 \usepackage[pdftex,
    pdffitwindow,
    plainpages=false,
    pdfpagelabels=true,
    pdfborder={0 0 0},
    bookmarks=true,
    bookmarksnumbered=true,
    bookmarksopen=false
 ]{hyperref}

\fi
\fi

% --- -----------------------------------------------------------------
% --- Document-specific definitions.
% --- -----------------------------------------------------------------
\newtheorem{definition}{Definition}

\newcommand{\concat}{{\,\|\,}}
\newcommand{\bits}{\{0,1\}}

% --- -----------------------------------------------------------------
% --- The document starts here.
% --- -----------------------------------------------------------------
\begin{document}
%\maketitle
\sloppy

\noindent Rutgers University\\
CS206: Introduction to Discrete Structures II, Spring 2013\\
Professor David Cash\\

\begin{center}
\LARGE{\textbf{Homework 1}}\\
\large{\textbf{\emph{Due at the beginning of class on Monday, Feb 4}}}
\end{center}

\vspace{.1in}

\noindent\textbf{Instructions:} Point values for each problem are listed.
Write your solutions neatly or type them up.  Typed solutions will also be
accepted via Sakai.

\begin{enumerate}


\item Let $n\geq 2$ and $A_1,\ldots,A_n$ be sets in some universe $S$.  In this
problem we will give a proof by induction of the identity
\[
\left(\bigcap_{i=1}^n A_i\right)^c = \bigcup_{i=1}^n A_i^c.
\]

\begin{enumerate}
\item (5 points)  State and prove the base case for an inductive proof, meaning
that the identity is true when $n=2$.

\item (5 points)  State and prove the inductive step, where one shows that the 
identity is true for general $n>2$, assuming it is true for $n-1$.
\end{enumerate}

\item Give sample spaces that model the outcomes for the following experiments.
You may use a regular expression or other formalisms that you find convenient.
(2 points each)
\begin{enumerate}

\item Rolling 3 dice.

\item Rolling a die until an even result comes up, or the die is rolled three
times.

\item Tossing a pair of coins until they both come up tails.  

\item Draw 2 balls from an urn which contains 6 balls, each with a
distinct label from $\{1,2,3,4,5,6\}$.

\item 
Draw 1 ball from the same urn, then replace it and draw a ball again.
\end{enumerate}


\item For each of the sample space, describe the events (as sets)
$A\cup B$ and $A \cap B$, when $A$ and $B$ are as follows.  (2 points each)
\begin{enumerate}

\item $A =$ ``$5$ is rolled exactly twice" and $B = $ ``dice values add to an
odd number".

\item $A =$ ``$1$ comes up exactly twice" and $B = $ ``$3$ comes up exactly
twice".

\item $A =$ ``both coins come up heads at the same time at some point" 
and $B = $ ``both coins come up tails at the same time at some point"

\item $A =$ ``$1$ is drawn at least once" and $B = $ ``$1$ is drawn twice".

\item $A =$ ``$1$ is drawn at least once" and $B = $ ``$1$ is drawn twice".

\end{enumerate}

\end{enumerate}



% --- -----------------------------------------------------------------

\bibliographystyle{plain}
\addcontentsline{toc}{section}{References}
\bibliography{conf,abbrev_short,crypto}



% --- -----------------------------------------------------------------
% --- The Appendix.
% --- -----------------------------------------------------------------

\begin{appendix}
\end{appendix}


\end{document}

