


\documentclass[11pt]{article}
\usepackage{fullpage}
\usepackage{amsthm}
\usepackage{amsthm,amsmath,amsfonts,amssymb,amstext}
\usepackage{latexsym,ifthen,url,rotating}
\usepackage[usenames,dvipsnames]{color}


% --- -----------------------------------------------------------------
% --- Document-specific definitions.
% --- -----------------------------------------------------------------
\newtheorem{definition}{Definition}

\newcommand{\concat}{{\,\|\,}}
\newcommand{\bits}{\{0,1\}}
\newcommand{\Range}{{\mathrm{Range}}}
\newcommand{\A}{{\mathcal{A}}}

% --- -----------------------------------------------------------------
% --- The document starts here.
% --- -----------------------------------------------------------------
\begin{document}
%\maketitle
\sloppy

\noindent Rutgers University\\
CS206: Introduction to Discrete Structures II, Spring 2013\\
Professor David Cash\\

\begin{center}
\LARGE{\textbf{Homework 5}}\\
\large{\textbf{\emph{Due at the beginning of class on Wednesday, March 27}}}
\end{center}

\vspace{.1in}

\noindent\textbf{Instructions:} Point values for each problem are listed.
Write your solutions neatly or type them up.  Typed solutions will also be
accepted via Sakai.

\begin{enumerate}


\item (10 points total) Two teams $x$ and $y$ are playing each other in the
World Series, which is a best-of-seven-game match that ends when one
team wins 4 games.  Assume that team $x$ wins each game with probability
$p$, and that the outcome of each game constitutes an independent trial.

\begin{enumerate}
\item (0.5 points) What is the probability that $x$ wins the first four games?

\item (2 points) What is the probability that $x$ wins four games after
at most five game have been played?

\item (2 points) What is the probability that $x$ will win four games before
$y$ wins four games?  (i.e., What is the probability that $x$ wins
the Series?)

\item (0.5 points) Calculate and simplify your answer in part (c) when $p=1/2$
and when $p=2/3$.

\item (1 point) Let $X$ be the random variable that counts the number of games
that are played.  What is $\Range(X)$?

\item (2 points) What is $P(X=7)$?

\item (2 points) What is $P(X\geq 6)$?
\end{enumerate}

\item (4 points) Suppose we roll two fair dice.  Let the random variable $X=$
``the minimum of the two dice'' and $Y =$ ``the absolute value of the
difference of the two dice''.  Find $E(X)$ and $E(Y)$.

\item (4 points) Suppose boxes of cereal are filled with a random prize,
each drawn from independently and uniformly from $6$ possible prizes.
If we buy $N$ boxes of cereal, what is the expected number of distinct
prizes we will collect? \begin{small}\textsf{Hint: Consider the indicator
random variables $I_{E_i}$ for the event $E_i =$ ``the $i$-th price was
in some box''.}\end{small}

\item (4 points) A group of $m$ men and $w$ randomly sit in a single row at a
theater.  If a man and woman are seated next to each other we say they form a
couple.  (Couples can overlap, meaning that one person can be a member of two
couples.)  What is the expected number of couples?
\begin{small}\textsf{Hint: Use indicator
random variables for each possible couple forming.
}\end{small}


\item (3 points) Suppose an experiment tosses a fair coin twice;  the experiment
``succeeds'' if both tosses were Heads.  We repeat this experiment 
for 12 independent trials.  Let $N$ be the random variable that counts
the fraction of trials that are successful (so $N = S/12$, where
$S$ is the number of successful trials).  Find $E(N)$.

\item \textbf{Extra Credit: (4 points)} Consider the experiment where $n$ balls
are to be placed randomly into $n$ boxes. Let $N_1$ count the number of boxes
with exactly one ball, and let $N_2$ count the number of boxes with exactly two
balls. Find the probability of the events ``$N_1 = n$" and ``$N_1 = n - 1$".
Use the indicator technique to find $E(N_1)$ and $E(N_2)$.


\end{enumerate}


\end{document}

